{
  \setlength{\parskip}{2pt}
  First of all, I would like to thank my first supervisor Harald Ita for his advice, guidance, fascinating insights
  and always giving me opportunities to grow as a researcher and giving me incentives to strive for higher standards.
  I am also very thankful to my second supervisor Fernando Febres Cordero for 
  being always ready to help me with all aspects of my doctoral studies,
  and letting me ask all the questions I ever wanted.
  I hope it was not too troublesome for you.

  I am very grateful to Ben Page and Samuel Abreu for patiently mentoring me,
  and very insightful discussions, research-related or otherwise.
  I cannot emphasize enough how valuable was it for me to have an opportunity to work by your side.

  I would like to thank my colleague Felix Anger, with whom I had worked closely in the first two years of my Ph.D.\ studies.
  It was a pleasure.

  My thanks go to Ekta Chaubey, Jerry Dormans, Michael Ruf, Wladimir Tschernow, Maximilian Klinkert, Harald Ita, Fernando Febres Cordero for critically reading parts of the manuscript and suggestions
  regarding the contents.

  I am also indebted to Ekta Chaubey for continuous encouragement, assistance, and for tolerating my complicated personality, especially during the time of writing this thesis. 

  I would like to thank Anna Gantimurova for supporting me in the first years of my doctoral studies.

  Finally, I would like to thank the ``Graduiertenkolleg 2044'' for providing a vibrant environment for carrying out Ph.D.\ research,   
  and for the unique opportunity of having a framework of dialogue with our colleagues from experiments. 

  %GRK for framework, seminars and events


  %%%%%%%

  %First and foremost I would like to thank my supervisor Harald Ita, for his scientific guidance and his supportive attitude. Your advice and help has been invaluable! I would also like to express my gratitude to my second supervisor Fernando Febres Cordero, for the good collaboration and his outstanding commitment. I really enjoyed working with you! 

  %I would like to thank my collaborators Stefan H{\"o}che and Daniel Ma{\^i}tre, for the things I have learnt from them.

  %I would like to thank my colleague and office mate Vasily Sotnikov, for all the discussions and good collaboration on many aspects of this work. We have been a good team! To all the current and former members of the 8\textsuperscript{th} (and 7\textsuperscript{th}) floor for supporting me in one or another way but also for the good times shared: Kicker tournaments, Christmas parties and the ``Betriebsausflug'' come to my mind. A special thanks goes to Samuel Abreu for reading the manuscript of this thesis, to Matthijs van der Wild for forming the winning team of the kicker tournament '17 and to Jerry Dormans for tolerating our long discussions in the office but also for the fun in ``office 809''. To the members of the $\hbar$-racing team for the shared memories on snow and beyond: Lukas Altenkamp, Michele Boggia, Michael Kordovan.

  %I would like to thank Hannah Arnold, Giulia Gonella and Gernot Knippen for the shared time as student representatives of the ``Graduiertenkolleg 2044''. From organizing BBQs to increasing the ``educational value'' of the seminar series: I think we provided some innovative impulses!
}
