The aim of this chapter is to give a brief overview of standard computational methods for the
evaluation of multi-loop scattering amplitudes, and establish notation at the same time.

\section{Dimensional Regularization}



As we discussed in \cref{sec:hadcoll}, loop amplitudes exhibit UV and IR divergences,
which must cancel in computations of any observable quantity.
Thus, we regularize divergences in intermediate stages and leave the cancellations to the end. 
In principle, any consistent regularization prescription should suffice.
However, certain features of a regularization method such as breaking of Poincaré invariance or gauge symmetries
can impact dramatically its computational efficiency.
For this reason, many different regularization schemes have been studied (for a recent review see \cite{Gnendiger:2017pys}).

In this thesis, we regularize divergences by analytically continuing the dimensionality of space-time from four to $D = 4 - 2\eps$ dimensions.
Divergences manifest themselves in the limit $D\to4$ (or $\eps\to0$) as poles in $\eps$.
This idea is known as \emph{dimensional regularization} (DR) \cite{tHooft:1972tcz}.
Dimensional regularization has several advantages over other regularization schemes.
It explicitly preserves both Poincaré and non-chiral gauge symmetries.\footnote{Chiral gauge symmetries are in general violated by DR and require special treatment.}
Furthermore, it can be used as IR and UV regulator at the same time.
This makes it by far the most practical regularization scheme.


DR regularizes $L$-loop integrals by carrying out the integrations in $D$ dimensions, i.e.\ the integration measure is replaced as
\begin{equation}
  \int \prod_{i=1}^{L} \frac{\dd[4]{\ell_i}}{(2\pi)^{4}} ~\longrightarrow~ \quantity(\mu^{4-D})^{L}~\int \prod_{i=1}^{L} \frac{\dd[D]{\ell_i}}{(2\pi)^{D}},
  \label{eq:dimregmeasure}
\end{equation}
where $\mu$ is an arbitrary mass scale introduced to preserve the dimensionality of the coupling constants.
Accordingly the algebra of the kinematic variables is extended to $D$ dimensions, implying for instance
\begin{equation}
   \pdv{\ell^\mu_i}{\ell_{j\,\mu}} = \delta_{i j} D.
\end{equation}

The fields and states transform as representations of Poincaré group,
implying that all Lorentz tensor and spinor structures (such as the metric tensor, momenta, external states, etc.\@) have
to be generalized to $D$ dimensions.
There are different flavors of DR making particular choices on how this generalization is accomplished (see e.g.\ \cite{Gnendiger:2017pys}).

The conceptually simplest variant of DR, the \emph{conventional dimensional regularization} (CDR)
treats all Lorentz structures uniformly in $D$ dimensions, effectively replacing the $\mathrm{SO}(1,3)$ Lorentz symmetry by $\mathrm{SO}(1,D-1)$.
CDR has been successfully employed in many analytic computations.
In this thesis, however, we are concerned with numerical methods, for which CDR is not practical due to
(formally infinite) number of additional external states.
Instead, we use the 't Hooft -- Veltman (HV) scheme which keeps all external states and momenta in four dimensions.
%and extends the space-time symmetry group to $\mathrm{SO}(1,3)\otimes \mathrm{SO}(D-4)$.

It is advantageous to distinguish in the intermediate stages of computation the dimensionality of loop integration (or loop momenta) $D$ and
the dimensionality of internal particles $D_s \geq  D$.
We will extensively exploit this separation in \cref{chap:dshel} to derive an efficient numerical algorithm
for the evaluation of amplitudes in the HV scheme.


\section{Loop Amplitudes}

For a particular choice of a gauge group, particle content, and operators of
a theory (which we collectively call a \emph{model}), Feynman rules can be derived.
Then the amplitudes in \cref{eq:amplitudes_expansion} can be obtained as a sum of all Feynman diagrams
with $N = 2+n$ specified external particles and $L$ loops.
For the purpose of this chapter, we assume that the amplitudes $A^{L}_N$ are Lorentz-scalars.
This can be achieved for example through decomposition into form factors (see \cref{chap:dshel}),
and details of this procedure are of no importance here.

The most general form for any $L$-loop amplitude with $N$ external particles with momenta $p_1,\ldots,p_N$ is
\begin{equation}
  \sum_{\Gamma \in \Delta} \sum_{\gamma_\Gamma} \qty(\int \prod_{l=1}^{L} \tilde{\dd{\ell_l}}) ~ 
  \frac{\mathcal{N}_{\Gamma,\gamma_\Gamma}%(\ell_1,\ldots,\ell_L,p_1,\ldots,p_{N-1})
    }{\prod_{j\in P_\Gamma} \rho_j^{\gamma_{\Gamma j}}},
    \qquad \tilde{\dd{\ell_l}} \coloneqq \frac{\dd[D]{\ell_l}}{(2\pi)^{D}}.
  \label{eq:general_amplitude}
\end{equation}
Here, $\rho_j = q_j^2 - m_j^2$ is a denominator of a propagator with momentum $q_j$ and mass $m_j$. The sum $\sum_{\gamma_\Gamma}$ runs through all assignments
of exponents $\gamma_{\Gamma j} > 0$ of each propagator $\rho_j$. 
A \emph{topology} $\Gamma$ is identified with a particular set of (unique) propagators $P_\Gamma$,
treating two sets of propagators equal if one can be transformed into the other with linear re-parametrizations of loop momenta.
Note that topologies defined in this sense might have multiple graph representations.
The external sum runs over a set of all distinct topologies $\Delta$ contributing to the amplitude.
The numerators $\mathcal{N}_{\Gamma,\gamma_\Gamma}$
are functions of all possible scalar products of loop momenta, external momenta and polarization states.

Carrying out loop integrations in \cref{eq:general_amplitude} is a challenging task in general.
Fortunately, there are many relations between integrals in \cref{eq:general_amplitude},
and it is thus sufficient to consider a significantly smaller set of integrals $\Delta_M \subset \Delta$, commonly referred to
as \emph{master integrals}.
Taking the integral relations into account the amplitude can be written in the form
\begin{equation}
  \sum_{\Gamma \in \Delta_M} \sum_{i\in M_\Gamma} ~ c_{\Gamma,i} ~ I_{\Gamma,i}, 
    \qquad I_{\Gamma,i} \coloneqq 
      \qty(\int \prod_{l=1}^{L} \tilde{\dd{\ell_l}}) \frac{m_{\Gamma,i}}{\prod_{j\in P_\Gamma} \rho_j^{\gamma_{\Gamma i,j}}},
  \label{eq:amplitude_integrated}
\end{equation}
where $M_{\Gamma}$ is a set of numerator insertions $m_{\Gamma,i}$ and exponents $\gamma_{\Gamma i}$.
Despite the fact that the numerator functions $\mathcal{N}_{\Gamma,\gamma_\Gamma}$ in \cref{eq:general_amplitude} depend on the types of external particles and a particular model,
the master integrals $I_{\Gamma,i}$ are universal and determined only by kinematics.

Evaluation of amplitudes can thus be separated into two different problems:
\begin{enumerate}
  \item Extracting  the list of all relevant topologies and numerator insertions from \cref{eq:general_amplitude},
    identifying the master integrals and computing them.
  \item Reducing all terms in \cref{eq:general_amplitude} to the form of \cref{eq:amplitude_integrated}, thus
    determining the coefficients $c_{\Gamma,i}$.
\end{enumerate}
The process- and model-dependent information is contained in the master-integral coefficients $c_{\Gamma,i}$,
so the master integrals can be recycled for all kinematically-equivalent amplitudes.

In this thesis, we focus on the reduction to master-integral coefficients.
We will review the standard approach in the next section, and
then explain how unitarity-based methods allow us to address some of its difficulties in \cref{chap:numunitarity}.

\section{Classification of Numerators}
\label{sec:classification_numerators}

To carry out the task of reduction,
we need to classify all possible structures that can appear in the numerator functions $\mathcal{N}_{\Gamma,\gamma_\Gamma}$ in \cref{eq:general_amplitude}.

We start by making a convenient choice of independent variables.
Due to Lorentz-invariance, the numerators $\mathcal{N}_{\Gamma,\gamma_\Gamma}$ can only depend on scalar products
\begin{equation}
  \label{eq:sps_all}
  \vb*{sp} = \{ \sp(\ell_i,\ell_j),~\sp(\ell_i,p_j),~\sp(\ell_i, \tau_j) \}, \qquad \dim{\vb*{sp}} = D_{\text{ext}} L + \frac{1}{2}L(L+1),
\end{equation}
where $D_{\text{ext}}$ is the dimensionality of external states and momenta.
Here the vectors $\tau_i$ are the basis of the orthogonal complement of $\spn{p_1,\cdots,p_{N-1}}$ to the full four-dimensional Minkowski space,
so we have $p_i \cdot \tau_j = 0$ and $\tau_i \cdot \tau_j = \delta_{ij}$.
The directions $\tau_i$ can be introduced into $\mathcal{N}_{\Gamma,\gamma_\Gamma}$ through external polarization states, hence,
the scalar products $\sp(\ell_i, \tau_j)$ will drop out if polarization sums are performed.
In the HV scheme, $D_{\text{ext}}=4$ and the number of corresponding scalar products are
\begin{equation}
  \label{eq:sp_counts}
  \renewcommand{\arraystretch}{1.5}
  \begin{tabular}[h]{p{10ex} p{20ex}}
    %$3N-10$ & $\sp(p_i,p_j)$ \\
       $\sp(\ell_i,\ell_j)$  &   $\dfrac{1}{2}L(L+1)$                \\
        $\sp(\ell_i,p_j)$    &   $\min\qty(4,N-1) ~L$                  \\
       $\sp(\ell_i, \tau_j)$ &   $ \qty\big(4-\min\qty(4,N-1)) L $  \\
  \end{tabular}
\end{equation}

The numerators $\mathcal{N}_{\Gamma,\gamma_\Gamma}$ can be thus represented as elements of a polynomial ring
$\mathbb{Q}_{\vb*{x},D}[\vb*{sp}]$ 
with coefficients $c_{\va{i}}(\vb*{x},D)$ take values int the field of rational functions $\mathbb{Q}_{\vb*{x},D}$ of external kinematics $\vb*{x}$
and $D$:~\footnote{
  $\vb*{x}$ is understood to be a list of variables corresponding to some parametrization of external kinematics, which we will not need to specify explicitly here.
}
\begin{equation}
  \mathcal{N}_{\Gamma,\gamma_\Gamma}(\vb*{x}, \vb*{sp}, D) = \sum_{\va{i} : \abs{\va{i}} < H_{\Gamma}} c_{\va{i}} (\vb*{x},D)~ m^{\va{i}}(\vb*{sp}),
  %m^{\va{i}}(\vb*{sp}) = (\sp(\ell_1,\ell_1))^{i_1} ~\cdots~ (\sp(\ell_1,p_1))^{i_{\min\qty(4,N-1)L}} ~\cdots~,
  \qquad m^{\va{i}}(\vb*{sp}) = \mathrm{sp}_1^{i_1} ~\ldots{}~ \mathrm{sp}_{\dim{\vb*{sp}}}^{i_{\dim{\vb*{sp}}}},
  %%(\sp(\ell_L, \tau_{4-\min\qty(4,N-1)}))^{\i_{\dim{\vb*{sp}}}}
\end{equation}
where the sum runs over all monomials $m^{\va{i}}(\vb*{sp})$ with \emph{total degree} $\abs{\va{i}} \equiv \sum_j i_j$ less than
some maximal value $ H_{\Gamma} $ determined by the UV structure of the model in consideration.


For a given topology $\Gamma$ in \cref{eq:general_amplitude} we chose $N_p$ denominators $\vb*{\rho}\coloneqq\{\rho_1,\ldots,\rho_{N_p}\} \in P_\Gamma$ as
independent variables. Then $N_p$ scalar products from the set $\vb*{sp}$
can be expressed as linear functions of denominators $\vb*{\rho}$.
This implies that whenever these scalar products are encountered in $\mathcal{N}_{\Gamma,\gamma_\Gamma}\qty(\vb*{sp})$,
they can be canceled with one of the denominators and attributed to a topology with fewer propagators,
leaving only a scalar remainder (i.e.\ independent of loop-momenta) in the numerator of $\Gamma$.
For this reason, these scalar products are commonly called \emph{reducible scalar products} (RSP).
We choose the remaining scalar products
\begin{equation}
  \vb*{\alpha} \coloneqq \vb*{sp} \setminus \{\text{RSPs}\} , \qquad N_\text{ISP} \coloneqq \dim{\vb*{\alpha}}   =  \dim{\vb*{sp}}-N_p,
\end{equation}
which are called
\emph{irreducible scalar products} (ISP) to be the remaining independent variables.

An important observation which becomes evident from our choice of variables is that all topologies with $N_p> 4 L +\frac{1}{2}L(L+1)$ are reducible.
Indeed, in this case we have $\dim\vb*{sp} < N_p$ so there are no irreducible scalar products, and only $(\dim\vb*{sp})$ denominators can be independent.

After a change of variables $\vb*{sp}\to \{\vb*{\rho},\vb*{\alpha}\}$ and cancellation of denominators
we find that the numerators are now polynomials in ISPs only:
\begin{equation} \label{eq:rsp_reduction}
  \mathcal{N}_{\Gamma,\gamma_\Gamma}(\vb*{x}, \vb*{sp}, D) \longrightarrow \qquad
    \mathcal{N}^{\prime}_{\Gamma,\gamma_\Gamma}(\vb*{x}, \vb*{\alpha}, D) =
    \sum_{\va{i} : \abs{\va{i}} < H_{\Gamma}} c^{\prime}_{\va{i}} (\vb*{x},D)~ m^{\va{i}}(\vb*{\alpha}),
\end{equation}
and all topologies with $N_p > 4L +\frac{1}{2}L(L+1)$ are eliminated.
In practice, this can be non-trivial and systematically performed with Gröbner basis techniques \cite{Zhang:2012ce,Mastrolia:2012wf,Mastrolia:2012an,Mastrolia:2016dhn}.
In unitarity-based approaches, this step is taken into account by construction,
and need not be carried out explicitly. We discuss this in more detail in \cref{sec:ansatz_integrand}

For $L=1$, the only irreducible scalar products are of the transverse type $\sp(\ell_i, \tau_j)$, which vanish upon integration due to Lorentz symmetry either directly,
or after some simple considerations (see \cref{sec:traceless_completion} for details).
For $L>1$, however, this is not the case and there are more identities that we can take advantage of to reduce the number of master integrals.
We discuss this in the next section.

We conclude this section with a remark that the variables $\{\vb*{\rho},\vb*{\alpha}\}$ we chose for parametrization of the integrand can be promoted
to the integration measure in \cref{eq:general_amplitude}. This representation of loop integrals known as the \emph{Baikov representation} \cite{Baikov:1996rk} 
makes $4 L +\frac{1}{2}L(L+1)$  non-trivial integrations manifest and allows to trivially integrate out irrelevant loop-momenta directions. This representation
exhibits many other useful properties, nonetheless, we will not need to adopt it in this thesis.


%We parametrize loop momenta as follows
%\begin{subequations}
  %\begin{equation}
    %\ell_l = \ell_{l[4]} + \ell_{l[D-4]}, \qquad 
    %\begin{array}[]{ll}
      %\ell^\mu_{l[4]} &\coloneqq g^{\mu\nu}_{[4]} \ell_\nu \\
      %\ell^\mu_{l[D-4]} &\coloneqq g^{\mu\nu}_{[D-4]} \ell_\nu \\
    %\end{array},
    %\qquad
    %\ell_{l[4]}\cdot\ell_{l[D-4]} = 0,
    %\label{eq:loop_momenta}
  %\end{equation}
  %and
  %\begin{align}
    %\ell_{l[4]} &=  \sum_{i\in \mathcal{B}_{\text{phys}}} c^{i}_l ~ v^{i} ~+~ \sum_{i\in \mathcal{M}_{[4]}\setminus\mathcal{B}_{\text{phys}}} \chi^{i}_l ~ \tau^i, \\ 
    %\ell_{l[D-4]} &=  \sum_{i\in\tilde{\mathcal{B}}} \mu^i_l ~ \tilde{n}^{i},
  %\end{align}
%\end{subequations}
%where $v^{i}$ are basis vectors of the vector space $\mathcal{B}_{\text{phys}} \coloneqq \spn{p_1,\cdots,p_{N-1}}$ spanned by $N-1$ external momenta;
%the vectors $\tau_i$ span the orthogonal complement of $\mathcal{B}_{\text{phys}}$ to the four-dimensional Minkowski space $\mathcal{M}_{[4]}$.


\section{Integration-by-Parts Identities}
\label{sec:ibp}


Using \cref{eq:rsp_reduction} we can bring \cref{eq:general_amplitude} to the form 
\begin{equation} \label{eq:rsp_reduced_amplitude}
  \sum_{\Gamma \in \Delta} \sum_{\gamma_\Gamma} \sum_{\va{i}} ~ c_{\Gamma\gamma_\Gamma, \va{i}} (\vb*{x},D) ~ 
  \qty(\int \prod_{l=1}^{L} \tilde{\dd{\ell_l}}) ~  
  \frac{m^{\va{i}}(\vb*{\alpha})%(\ell_1,\ldots,\ell_L,p_1,\ldots,p_{N-1})
    }{\prod_{j\in P_\Gamma} \rho_j^{\gamma_{\Gamma j}}}.
\end{equation}
The integrals in \cref{eq:rsp_reduced_amplitude} are process-independent and have the desired form of \cref{eq:amplitude_integrated}.
However, we are still missing a large number of identities, which are present due to the vanishing of the total derivative in dimensional regularization.
These identities are known as \emph{integration-by-parts} (IBP) identities \cite{Chetyrkin:1981qh,Tkachov:1981wb} and have a general
form of
\begin{equation}
  \qty(\int \prod_{l=1}^{L} \tilde{\dd{\ell_l}}) ~  \pdv{\ell_k^\mu}(
  \frac{ v_k^\mu \ommit{(\vb*{x},\vb*{\rho},\vb*{\alpha},D)} ~ \, m^{\va{i}}(\vb*{\alpha})
    }{\prod_{j} \rho_j^{\gamma_{j}}}
    ) = 0,
  \label{eq:ibps}
\end{equation}
where the components of \emph{IBP vectors} $v_k^\mu(\vb*{x},\vb*{\rho},\vb*{\alpha},D)$ are arbitrary rational functions
of all variables. The monomials $m^{\va{i}}(\vb*{\alpha})$ can be absorbed in $v_k^\mu$, but
it is convenient to keep them separate.

It is useful to organize all topologies $\Delta$ in \cref{eq:rsp_reduced_amplitude}
hierarchically by defining a partial ordering
\begin{equation} \label{eq:topology_order}
    \Gamma_1 > \Gamma_2 \iff P_{\Gamma_1} \supset P_{\Gamma_2},
\end{equation}
meaning that topology $\Gamma_2$ is obtained from topology $\Gamma_1$ by removing one or more propagators.
We then slice $\Delta$ into strictly ordered (overlapping) subsets $\Delta_i$ such that
\begin{subequations} \label{eq:topology_sequence}
    \begin{gather}
         \Delta = \bigcup_i \Delta_i,  \\
         \Delta_i  = \{\Gamma \in \Delta : \forall \, \Gamma_1, \Gamma_2 \in\Delta \qif\qty( \Gamma_1 \in \Delta_i \qand \Gamma_2 < \Gamma_1) \qthen \Gamma_2 \in \Delta_i \}.
    \end{gather}
\end{subequations}
A subset $\Delta_i$ can thus be identified by its ``greatest'' topology $\Gamma_{\Delta_i}$ ---
the topology with the maximal number of propagators, which we will call a \emph{parent} topology (or diagram).

If we allow exponents $\gamma_{\Gamma j}$ in \cref{eq:rsp_reduced_amplitude} to also be $\leq0$,
all integrals in \cref{eq:rsp_reduced_amplitude} can be cast into \emph{families}
\begin{equation} \label{eq:integral_families}
    I_{\Delta_i}[\va{\gamma},\va{i}] \coloneqq 
  \qty(\int \prod_{l=1}^{L} \tilde{\dd{\ell_l}}) ~  
  \frac{m^{\va{i}}_{\Gamma_{\Delta_i}}(\vb*{\alpha})%(\ell_1,\ldots,\ell_L,p_1,\ldots,p_{N-1})
    }{\prod_{j\in P_{\Gamma_{\Delta_i}}} \rho_j^{\gamma_{j}}}.
\end{equation}
IBP identities of \cref{eq:ibps} generate an under-determined system of homogeneous linear equations between integrals inside a family.\footnote{
  In some exceptional cases IBP identities can relate integrals between different families, but we will not need to consider them in this thesis. 
}
The integrals which cannot be determined from the system are chosen to be master integrals.
The remaining integrals are then expressed as linear combinations of master integrals.
A systematic approach of generating and solving a complete system of IBP identities was formulated by Laporta \cite{Laporta:2001dd}.
There are multiple publicly-available implementations such as
\textsc{Reduze} \cite{Studerus:2009ye,vonManteuffel:2012np},
\textsc{Fire} \cite{Smirnov:2008iw,Smirnov:2014hma},
\textsc{LiteRed} \cite{Lee:2012cn,Lee:2013mka},
\textsc{Kira} \cite{Maierhoefer:2017hyi,Maierhofer:2018gpa}, etc.

The task of IBP reduction is in general very difficult.
This is due to a large number of relations produced by \cref{eq:ibps},
which is in turn due to the number of choices of IBP vectors, as well as the fact that IBP identities,
in general, include the relations for integrals with raised denominator powers, even
though these integrals are mostly not found from Feynman rules.
In the end, the bottleneck is coming from the requirement to solve large systems of linear equations.
Many great improvements to the IBP-reduction technology have been suggested
\cite{Gluza:2010ws,Schabinger:2011dz,Larsen:2015ped,Bern:2017gdk,Chawdhry:2018awn,Kosower:2018obg,Mastrolia:2018uzb,Frellesvig:2019kgj,Frellesvig:2019uqt,Bendle2019}
since the Laporta algorithm became widespread.
It was also proposed to reconstruct the reduction tables from
numerical evaluations over finite fields \cite{vonManteuffel:2014ixa,Peraro:2016wsq,Peraro:2019svx,Klappert:2019emp,Smirnov:2019qkx}.
However, there is still a problem
that even when the analytic reduction tables can be obtained,
it might be challenging to use them for the reduction of amplitudes due
to the prohibitive size of expressions (see e.g.\ \cite{Chawdhry:2018awn,Bendle2019}).
Because of these complications it seems beneficial to carry out IBP reduction numerically and reconstruct
analytic expressions for amplitudes instead \cite{Badger:2018enw,Chicherin:2018yne,Badger:2019djh}.

In this thesis, we will take advantage of generalized unitarity-based methods (see \cref{chap:numunitarity}),
which do not require solving IBP systems and thus,
avoid some of the problems discussed above.

\section{Master Integrals}

Once the set of master integrals is found, the IBP reduction relations can be applied 
to all terms in \cref{eq:rsp_reduced_amplitude}.
This determines the coefficients of master integrals
\begin{equation} \label{eq:master_integrals}
    \qquad I_{\Gamma,i} = 
    \qty(\int \prod_{i=1}^{L} \tilde{\dd{\ell_i}}) \frac{m_{\Gamma,i}}{\prod_{j\in P_\Gamma} \rho_j^{\gamma_{\Gamma i,j}}}.
\end{equation}
in \cref{eq:amplitude_integrated}.

A variety of methods have been developed for the evaluation of the master integrals.
Broadly speaking, they can be classified into analytic and numerical methods.

Direct integration using a convenient choice of integration variables is the most straightforward option.
Most one-loop integrals have been obtained this way \cite{vanHameren:2010cp,Ellis:2007qk,tHooft:1978jhc,Denner:2010tr}.
However, for multi-scale multi-loop integrals this becomes  cumbersome   very quickly.
Another method is the integration in the Mellin-Barnes  representation  \cite{Smirnov:1999gc,Tausk:1999vh,Dubovyk:2017cqw}, which
has been used to solve some two-loop integrals.
However, the most commonly used and successful method is to obtain analytic results via
differential equations \cite{Kotikov:1990kg,Remiddi:1997ny,Gehrmann:1999as,Henn:2013pwa,Argeri:2007up,Henn:2014qga}
brought into their canonical form \cite{Henn:2013pwa}.
These equations are obtained by noticing that
derivatives of master integrals with respect to
external invariants or masses can be IBP-reduced to the linear vector space of master integrals from the same family.
Thus the whole family of integrals can be obtained from solving the system of linear differential equations.

On the numerical side (see e.g.\ \cite{Freitas:2016sty} for a comprehensive review),
the commonly used general methods are sector decomposition \cite{Binoth:2000ps,Binoth:2003ak}
and numerical integration in Mellin-Barnes representation \cite{Czakon:2005rk,Anastasiou:2005cb}.
Usually, the knowledge of analytical expressions is preferable since it allows for better control over
precision and performance required for Monte-Carlo integration over external phase space. However,
some notable NNLO QCD computations have been carried out with at least one of the integrals evaluated through numerical
methods \cite{Jones:2018hbb,Grazzini:2018bsd,Baglio:2018lrj}.

There are no constraints on the choice of basis of master integrals, apart from the linear-independence under IBP relations. 
This freedom can be exploited to facilitate particular aspects of the computation.
The basis of uniformly transcendental (UT) or pure integrals \cite{ArkaniHamed:2010gh} has a simple form of the
associated differential equations \cite{Henn:2013pwa}, which allows to solve them in terms of iterated integrals.
For numerical evaluations, as well as for exploring the structure of divergences of amplitudes,
it is often beneficial to make master integrals finite by constructing the so-called local integrands \cite{ArkaniHamed:2010kv,ArkaniHamed:2010gh,Badger:2016ozq,Badger:2016egz}.
Another numerically-oriented choice is the quasi-finite basis \cite{vonManteuffel:2014qoa,Panzer:2014gra}.
Finally, in the numerical unitarity method, which we discuss in \cref{chap:numunitarity}, the master integrands are required to have no doubled propagators.
The conversion between different master-integral bases is often simpler than obtaining the relations for high-degree numerator insertions,
and can be performed with the standard IBP reduction techniques.

In this thesis, we will not consider the evaluation of master integrals.
For the class of process we are interested in, the analytic results are available in the literature \cite{Gehrmann:2018yef,Papadopoulos:2015jft,Gehrmann:2000zt}.
